### 2025-08-06 actuarial.qmd
20250806 - additions

Hundreds of studies have examined clinical versus actuarial prediction methods across many disciplines.
Findings consistently show that actuarial methods are as [accurate](#sec-validity) or more [accurate](#sec-validity) than human judgment/prediction methods.
"There is no controversy in social science that shows such a large body of qualitatively diverse studies coming out so uniformly...as this one" [@Meehl1986, pp. 373–374].

>>>>

Hundreds of studies have examined clinical versus actuarial prediction methods across many disciplines.
Findings consistently show that actuarial methods are as [accurate](#sec-validity) or more [accurate](#sec-validity) than human judgment/prediction methods [@Dawes1989; @Aegisdottir2006; @Baird2000; @Grove1996; @Grove2000].
"There is no controversy in social science that shows such a large body of qualitatively diverse studies coming out so uniformly...as this one" [@Meehl1986, pp. 373–374].
In general, actuarial methods tend to be about 10% more accurate than human judgment [@Grove2000].
Thus, actuarial methods have [incremental validity](#sec-incrementalValidity) over human judgment.
As noted earlier, the superiority of actuarial approaches to clinical judgment tends to hold even when the clinician is given more information than the actuarial approach [@Dawes1989].
In addition, actuarial predictions outperform human judgment even when the human is given the result of the actuarial prediction [@Kahneman2011].
Allowing clinicians to override actuarial predictions consistently leads to lower predictive accuracy [@Garb2019].
